---
title: "Homework 3"
author: "Yuki Joyama"
date: "2023-10-05"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)

library(tidyverse)
```

# Problem 1

**Data import setup**
```{r}
library(p8105.datasets)
data("instacart") # call instacart data from the library 
```

The instacart data has `r nrow(instacart)` rows and `r ncol(instacart)` columns. Some key variables include order_id, product_id, user_id, product_name, etc.

**Examples of observations in instacart dataset**
```{r}
head(instacart, n = 8) |> 
  knitr::kable(digits = 1) # include table in the md file
```

The above example shows detailed information about `order_id` = 1. As we can see from the table, `order_id` is tied to specific IDs which identify product, user, aisle and department. From `product_name`, we can tell what the user got for this order. `order_number` = 4 means that this was the 4th order from this customer. `order_dow`, `order_hour_of_day`, and `days_since_prior` tell us that this order was placed on the 4th week, between 10 am to 11 am, and 9 days since the last order. `reordered` is recorded as 1 if the product has been ordered by this user in the past, and 0 otherwise. 

```{r}
# summarize aisle data
df_aisles <- instacart |> 
  group_by(aisle_id, aisle) |> 
  summarise(count = n())  # count how many times aisle_id were recorded in the dataset for each id (= the number of items ordered from each aisle)

max_aisle <- max(pull(df_aisles, count)) # check maximum count of aisles

df_aisles_max <- df_aisles |> 
  filter(count == max_aisle) # filter by the maximum count of aisles
```

`r max(pull(instacart, aisle_id))` aisles are in this dataset, and **`r pull(df_aisles_max, aisle)`** are the most items (in total `r max_aisle` items) ordered from.

The following plot shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered.
```{r}
# make a plot (the number of items ordered in each aisle)
df_aisles |> 
  filter(count > 10000) |> # just show aisles with more than 10000 items 
  ggplot(aes(x = reorder(aisle, count), y = count)) + 
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(
    title = "Items by aisle",
    x = "Aisle",
    y = "Item count",
    caption = "Only showing aisles with more than 10000 items ordered; sorted in an ascending order"
  )
```

**Popular items in "baking ingredients"**
```{r}
# table showing the three most popular items in "baking ingredients"
df_baking_ingredients <- instacart |> 
  filter(aisle == c("baking ingredients")) |> # filter by aisle
  group_by(product_name) |> 
  summarise(count = n()) |> # add count for each item
  arrange(desc(count)) |> # arrange in descending order
  head(n = 3) # only show the first three items
  
df_baking_ingredients |> 
  knitr::kable(digits = 1) # show table in md file 
```
The three most popular items in "baking ingredients" are `r pull(df_baking_ingredients, product_name)`.

**Popular items in "dog food care"**
```{r}
# table showing the three most popular items in "dog food care"
df_dog_food_care <- instacart |> 
  filter(aisle == c("dog food care")) |> # filter by aisle
  group_by(product_name) |> 
  summarise(count = n()) |> # add count for each item
  arrange(desc(count)) |> # arrange in descending order
  head(n = 3) # only show the first three items
  
df_dog_food_care |> 
  knitr::kable(digits = 1) # show table in md file 
```
The three most popular items in "dog food care" are `r pull(df_dog_food_care, product_name)`.

**Popular items in "packaged vegetables fruits"**
```{r}
# table showing the three most popular items in "packaged vegetables fruits"
df_packaged_vegetables_fruits <- instacart |> 
  filter(aisle == c("packaged vegetables fruits")) |> # filter by aisle
  group_by(product_name) |> 
  summarise(count = n()) |> # add count for each item
  arrange(desc(count)) |> # arrange in descending order
  head(n = 3) # only show the first three items
  
df_packaged_vegetables_fruits |> 
  knitr::kable(digits = 1) # show table in md file 
```
The three most popular items in "packaged vegetables fruits" are `r pull(df_packaged_vegetables_fruits, product_name)`.

The table below provides the mean order time for Pink Lady Apples and Coffee Ice Cream on each day of the week:

**Pink Lady Apples**
```{r warning = FALSE}
# Pink Lady Apples and Coffee Ice Cream table
pink_lady_apples <- instacart |> 
  filter(product_name == c("Pink Lady Apples")) |> # filter by product name
  select(-product_name) |> 
  group_by(order_dow) |> 
  summarise("mean_hour" = mean(order_hour_of_day)) # show the mean hour of the day on each day of the week

pink_lady_apples |> 
  knitr::kable(digits = 1) # show table in md file 
```

**Coffee Ice Cream**
```{r warning = FALSE}
# Pink Lady Apples and Coffee Ice Cream table
coffee_ice_cream <- instacart |> 
  filter(product_name == c("Coffee Ice Cream")) |> # filter by product name
  select(-product_name) |> 
  group_by(order_dow) |> 
  summarise("mean_hour" = mean(order_hour_of_day)) # show the mean hour of the day on each day of the week


coffee_ice_cream |> 
  knitr::kable(digits = 1) # show table in md file 
```

# Problem 2

**Data import setup**
```{r}
library(p8105.datasets)
data("brfss_smart2010") # call BRFSS data from the library 
```

**Data cleaning**
```{r warning = F}
df_brfss_cleaned <- brfss_smart2010 |> 
  janitor::clean_names() |> 
  filter(topic == "Overall Health" & response == c("Excellent", "Poor")) |>  # focus on Overall Health; only include responses from "Excellent" to "Poor"
  mutate(
    response = factor(response, levels = c("Poor", "Excellent")) # change response as factor taking levels ordered from “Poor” to “Excellent”
  )
```

















