---
title: "Homework 3"
author: "Yuki Joyama"
date: "2023-10-05"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)

library(tidyverse)
```

# Problem 1

**Data import setup**
```{r}
library(p8105.datasets)
data("instacart") # call instacart data from the library 
```

The instacart data has `r nrow(instacart)` rows and `r ncol(instacart)` columns. Some key variables include order_id, product_id, user_id, product_name, etc.

**Examples of observations in instacart dataset**
```{r}
head(instacart, n = 8) |> 
  knitr::kable(digits = 1)
```

The above example shows detailed information about `order_id` = 1. As we can see from the table, `order_id` is tied to specific IDs which identify product, user, aisle and department. From `product_name`, we can tell what the user got for this order. `order_number` = 4 means that this was the 4th order from this customer. `order_dow`, `order_hour_of_day`, and `days_since_prior` tell us that this order was placed on the 4th week, between 10 am to 11 am, and 9 days since the last order. `reordered` is recorded as 1 if the product has been ordered by this user in the past, and 0 otherwise. 











